---
title: "Language Education Data"
execute: 
  echo: false
---

```{r}
#| message: false
#| label: setup

library(tidycensus)
library(tidyverse)
```


```{r}
#| label: data
#| message: false


# Find and display correct variables from the S1603 table
variables <- load_variables(2022, "acs5", cache = TRUE) %>%
  filter(str_detect(name, "S1603"))

# Define the correct variables based on the output of load_variables
variables <- c(
  english_total = "S1603_C02_001E",   # Total population speaking only English
  spanish_total = "S1603_C06_001E",   # Total population speaking Spanish
  other_total = "S1603_C04_001E",     # Total population speaking other languages
  english_bachelors = "S1603_C02_016E", # Bachelor's degree or higher for English speakers
  spanish_bachelors = "S1603_C06_016E", # Bachelor's degree or higher for Spanish speakers
  other_bachelors = "S1603_C04_016E"    # Bachelor's degree or higher for other language speakers
)

# Fetch data for the entire US
data <- get_acs(
  geography = "us",
  variables = variables,
  year = 2022,
  survey = "acs5",
  output = "wide"
)

# Calculate percentage of population with bachelor's degree
data <- data %>%
  mutate(
    english_percentage_bachelors = (english_bachelors / english_total) * 100,
    spanish_percentage_bachelors = (spanish_bachelors / spanish_total) * 100,
    other_percentage_bachelors = (other_bachelors / other_total) * 100
  )

# Reshape data for plotting
data_long <- data %>%
  select(english_percentage_bachelors, spanish_percentage_bachelors, other_percentage_bachelors) %>%
  pivot_longer(cols = everything(), names_to = "language", values_to = "percentage_bachelors") %>%
  mutate(language = case_when(
    language == "english_percentage_bachelors" ~ "English",
    language == "spanish_percentage_bachelors" ~ "Spanish",
    language == "other_percentage_bachelors" ~ "Other"
  ))

# Create the bar graph using ggplot2

```

```{r}
#| label: plot

ggplot(data_long, aes(x = language, y = percentage_bachelors)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.7) +
  labs(
    title = "Percentage of Population with Bachelor's Degree by Language Spoken at Home",
    subtitle = "In the US, most Non-English Speakers are almost as likely to get bachelors degrees as\nEnglish Speakers",
    x = "Language Spoken at Home",
    y = "Percentage with Bachelor's Degree"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Using US census data to derive how the language that you speak in the US is correlated to your educational status revealed some interesting discoveries. It is quite interesting to see how you are almost just as likely to get a bachelors degree as someone who speaks Does not speak English or Spanish in the household as someone who only speaks English. It is unfortunate to see Spanish Speakers lacking behind. As a person who speaks Spanish in the house myself it is disheartening but also motivating to see the massive gap between bachelor rates of Spanish speakers and those of English speakers.

