---
title: "Language Education Data"
execute: 
  echo: false
---

```{r}
#| message: false
#| label: setup

library(tidycensus)
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```


```{r}
#| label: data
#| message: false

# Define the years of interest
years <- 2015:2022

# Function to fetch and process data for a specific year
fetch_process_data <- function(year) {
  # Find and display correct variables from the S1603 table
  variables <- load_variables(year, "acs5", cache = TRUE) %>%
    filter(str_detect(name, "S1603"))

  # Define the correct variables based on the output of load_variables
  variables <- c(
    english_all = "S1603_C02_001E",
    spanish_all = "S1603_C06_001E",
    other_all = "S1603_C04_001E",
    english_youth = "S1603_C02_002E",
    spanish_youth = "S1603_C06_002E",
    other_youth = "S1603_C04_002E",
    english_bachelors = "S1603_C02_016E",
    spanish_bachelors = "S1603_C06_016E",
    other_bachelors = "S1603_C04_016E",
    english_grads = "S1603_C02_015E",
    spanish_grads = "S1603_C06_015E",
    other_grads = "S1603_C04_015E",
    english_hs = "S1603_C02_014E",
    spanish_hs = "S1603_C06_014E",
    other_hs = "S1603_C04_014E"
  )

  # Fetch data for the entire US
  data <- get_acs(
    geography = "us",
    variables = variables,
    year = year,
    survey = "acs5",
    output = "wide"
  )

  # Calculate derived variables
  data <- data %>%
    mutate(
      english_total = english_all - english_youth,
      spanish_total = spanish_all - spanish_youth,
      other_total = other_all - other_youth,
      english_percentage_hs = ((english_bachelors + english_grads + english_hs) / english_total) * 100,
      spanish_percentage_hs = ((spanish_bachelors + spanish_grads + spanish_hs) / spanish_total) * 100,
      other_percentage_hs = ((other_bachelors + other_grads + other_hs) / other_total) * 100,
      english_percentage_bachelors = (english_bachelors / english_total) * 100,
      spanish_percentage_bachelors = (spanish_bachelors / spanish_total) * 100,
      other_percentage_bachelors = (other_bachelors / other_total) * 100,
      english_percentage_grads = ((english_bachelors + english_grads) / english_total) * 100,
      spanish_percentage_grads = ((spanish_bachelors + spanish_grads) / spanish_total) * 100,
      other_percentage_grads = ((other_bachelors + other_grads) / other_total) * 100
    )

  # Add a year column for later aggregation
  data <- data %>%
    mutate(year = year)

  return(data)
}

# Fetch and process data for all years
data_all_years <- map_dfr(years, fetch_process_data)

# Reshape data for plotting
data_long <- data_all_years %>%
  select(year, english_percentage_bachelors, spanish_percentage_bachelors, other_percentage_bachelors, 
         english_percentage_hs, spanish_percentage_hs, other_percentage_hs, 
         english_percentage_grads, spanish_percentage_grads, other_percentage_grads) %>%
  pivot_longer(
    cols = -year,
    names_to = c("language", "education_level"),
    names_pattern = "(.*)_(percentage_.*)",
    values_to = "percentage"
  ) %>%
  mutate(language = case_when(
    language == "english" ~ "English",
    language == "spanish" ~ "Spanish",
    language == "other" ~ "Other"
  ))

# Calculate mean percentages for each language and education level across years
mean_data <- data_long %>%
  group_by(year, education_level, language) %>%
  summarise(mean_percentage = mean(percentage, na.rm = TRUE))

```

```{r}
#| label: plot

ggplot(mean_data, aes(x = year, y = mean_percentage, color = language, linetype = education_level)) +
  geom_line(linewidth = 1) +
  labs(title = "Mean Percentage by Education Level and Language (2015-2022)",
       subtitle = "Education rates vary wildly but all seem to increase at the same rate",
       x = "Year",
       y = "Mean Percentage",
       color = "Language",
       linetype = "Education Level",
       caption = "Source: US Census Bureau") +
  theme_minimal()

```

21.6% of Americans speak a language other than English at home. The following data was pulled from the US Census Bureau. The relevant variables were pulled to analyse the relationship between language spoken at home and education rate. While Spanish speakers seem to be at a disadvantage when it comes to education, speakers of other languages are doing almost as well as English speakers.

